---
title: "splots: visualization of data from assays in microtitre plate or slide format"
author: "Wolfgang Huber"
package: splots
date: "2021-01-04"
output: 
    BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{splots: visualization of data from assays in microtitre plate or slide format}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.dim=c(6, 8))
```

This package is provided to support legacy code and reverse dependencies, but it should not be used as a dependency for new code development. It provides a single function, `plotScreen`, for visualising data in microtitre plate or slide format. As a better alternative for such functionality, please consider the `r CRANpkg("platetools")` package (see also https://github.com/Swarchal/platetools), or generic `r CRANpkg("ggplot2")` graphics functionality. 
  
Here we show a short demo of its parameters. First, we generate example data:

```{r exampledata} 
library("splots")
screen = lapply(1:12, function(i) cumsum(rnorm(384)))
names(screen) = paste("plate", LETTERS[seq(along=screen)])
```

```{r fig1}
plotScreen(screen, ncol=3)
```

```{r fig2}
plotScreen(screen, ncol=2, do.legend=FALSE)
```

```{r fig3}
plotScreen(screen, ncol=4, fill=c("white", "darkblue"), 
           main="Example data", 
           legend.label=expression(bar(x) == sum(x[i], i==1, n)))
```

```{r fig4}
for(i in seq(along=screen))
  screen[[i]][sample(384, 5)] = NA
plotScreen(screen, ncol=4, do.names=FALSE, 
           main="Example data", legend.label="Legend label")
```


```{r sessionInfo}
sessionInfo()
``` 
